Data Capsules: A Cloud Service for Secure Remote Access to Sensitive Data

### 2014-07-30 ccsw14/paper.tex
many changes; late draft of paper

\title{Improving Usability of Remote Access to Sensitive Data}

\author{
Draft Copy: \today\ \currenttime
}

%\numberofauthors{2} 
%
%\author{
%\alignauthor Alexander Crowell \\
%       \affaddr{Computer Science and Engineering Division}\\
%       \affaddr{University of Mighigan}\\
%       \affaddr{Ann Arbor, MI, USA}\\
%       \email{crowella@umich.edu}
%\alignauthor Atul Prakash \\
%       \affaddr{Computer Science and Engineering Division}\\
%       \affaddr{University of Mighigan}\\
%       \affaddr{Ann Arbor, MI, USA}\\
%       \email{aprakash@umich.edu}
%}

@@

In this paper, we propose Data Capsules, a system that is designed to enable
access to sensitive data for analysis by trusted remote users, while
maintaining reasonable guarantees of data security.  Data Capsules uses
virtualization to provide remote users with a privileged, but secure
environment into which they can bring arbitrary, and even potentially
malicious, software or data in order to analyze sensitive data, while
minimizing the available channels for data leaks.  Our early implementation
realizes much of this protection, providing a basic framework for secure
analysis of data that addresses many aspects of network, storage, and covert
channel security.

@@

\keywords{data capsules, data privacy}

@@

In the age of the Internet, new data is being created at incredible rates
\cite{digital-universe}.  Along with these waves of new data, new tools have
in turn been created to help analyze and gain insight from it. Unfortunately,
access to data is not always a simple matter.  Some types of data may be
sensitive, including data protected by copyright, data representing the personal
information of individuals, as well as any data that may contain secrets that
its owner does not wish to have revealed, for example corporate or state secrets.

The current state-of-the-art \cite{vrdc} in systems developed for the purpose
of enabling access to these types of data use remotely accessed machines, which
can then be used by researchers without the need for travel to conduct their
analysis. But beyond this apparent convenience, rules for using such systems
are often highly restrictive, with researchers limited to using predefined
software environments, and being subject to strict auditing requirements for
all software and data that is uploaded to these software environments.

To solve this substantial inconvenience, we propose a new system which we call
Data Capsules.  While, to an end user, a Data Capsules environment will
minimally differ from the type of environment provided by the CMS Virtual
Research Data Center, its design grants end users privileged access to install
software of their choice --- even down to the virtualization-compatible
operating system of their choice --- without requiring a review process and
without any compromise to the security of the data being provided.

Data Capsules makes use of virtualization as a key element in achieving this
security and usability goal.  Under the Data Capsules use model, a researcher
would be allocated a virtual machine (henceforth abbreviated as VM, or guest
VM), which they could then prepare and maintain with whatever software and
underlying operating system is best suited to the intended purpose.  Each Data
Capsules VM is endowed with two modes of operation.  In \emph{maintenance} mode,
the researcher is endowed with full Internet access, which enables them to
easily configure and maintain the software environment almost as if they were
using a their own personal machine; however, in maintenance mode, the researcher
does not have access to the sensitive data.  In order to gain access to this
data, the user must switch their VM to the second mode, which we call
\emph{secure} mode.  In secure mode, the user's VM state is saved, and
execution continues on an alternate machine that is firewalled to prevent
Internet access, but which does have access to the sensitive data.

@@

When a user wishes to further change their software environment in order to
continue data analysis, they may then switch back to maintenance mode, after
which the secure storage volume will detatch, and all VM state since the
previous switch to secure mode is discarded, with the VM continuing to run from
the point it left off at when last in maintenance mode.  In this way, every
trace of the operations performed on the sensitive data is erased and forgotten.

Furthermore, assuming any sensitive data is read-only, and is properly backed
up, Data Capsules functionality means that there is no need to audit installed
software for malware, as even infected software would be unable to exfiltrate
the data from the secure mode environment.

@@

While this system provides access to Census data, and seems robust to data
leaks, it is also highly restrictive, requiring researchers to relocate and use
foreign hardware and potentially software as well, to conduct their research.

@@

However, this state-of-the-art solution still suffers from significant
drawbacks, in that researchers are largely limited to using the preallocated
environments supplied by the VRDCs.  In order to upload software or data not
already available in a virtual desktop, it must first undergo a manual review
process.  For large software suites, this can at best add significant
inconvenience, and at worst expose the system to a malicious attack, as there is
no method for reliably detecting malicious software.  Data Capsules is designed
so that no review process is necessary to upload software or data to an
environment, while providing strong guarantees of data privacy by design.

@@

QubesOS \cite{qubes} is a system for isolating a user's desktop environment into
secure compartments, called ``Qubes," which are intended to prevent malicious
code from compromising the entire system by restricting its effect to the
compartment in which it resides.  Although Qubes is very effective for this 
purpose, it would not serve as a particularly secure tool for data access, as it
suffers from the same challenge of how to move data from one compartment, which
is presumably unrestricted, into another compartment containing the sensitive
data to be operated on.  Data Capsules offers a clear design for allowing the
import of data and code into an environment that can still offer secure access
to sensitive data.

Storage Capsules \cite{capsules-borders} is a virtualization-based system for
protecting data privacy on individual machines.  Under Storage Capsules, a
machine runs a virtual machine manager (VMM) system, which in turn runs the
user's operating system environment within a VM.  On the same machine, there is
a second VM running an interface that serves as a ``storage capsule" for
sensitive data.  The user's main VM is endowed with two modes of operation,
\emph{normal} and \emph{secure} mode, and similar to Data Capsules, normal mode
grants full network priviliges while secure mode grants access to the sensitive
data, via communication with the storage capsule VM.

@@

Specialized operating systems that enforce information flow control, such as
HiStar \cite{histar}, offer low-level operating system primitives that allow
developers to design programs in such a way that data leaks are not possible.
As with SELinux, there is still a substantial burden of having to craft policies
that prevent these malicious data flows.  Additionally, programs must be
specifically developed for compatibility with HiStar, creating an even greater
burden on users so long as that system is not widely supported by the software
community.

@@

@@

The major element that is used to isolate VMs that are in the middle of
accessing sensitive data is live migration \cite{migration-clark}.  When a VM
switches from maintenance mode to secure mode, the underlying virtual machine
manager migrates that VM to a separate physical machine that has access to the
sensitive data, but which is configured to lack network access for any purpose
other than to receive other incoming VM migrations.  In other words, VMs in
maintenance mode are strictly segregated within separate physical machines from
VMs running in secure mode.  We can therefore refer to a physical machine
running only maintenance VMs as a \emph{maintenance machine}, and likewise a
\emph{secure machine} runs only secure mode instantiations of VMs.  We illustrate an example deployment in Figure \ref{fig:configuration}.

@@

\caption{An example deployment configuration for a Data Capsules system.
Dotted boxes denote physical separation, which is necessary between maintenance
machines, secure machines, and the remote access proxy.  The sensitive data
server may exist separately, or it may reside on one of the secure machines.
The only connections between maintenance and secure machines will be for the
purpose of tranferring virtual machine state to secure machines, and for
read-only accesses to the VM disk images.  This data largely travels in one
direction, from maintenance machines to secure machines.}

@@

By adopting this segregated design, the threat of data leak via covert channels
within a single physical machine is eliminated, so long as the trusted computing
base remains uncompromised.  The remaining risk is that of one VM in secure mode
leaking data to another VM that is also in secure mode, but there is no reason
to want to do so, as both VMs would already have access to the entire secure
repository.  On a machine running VMs in maintenance mode, there is no way to
access the sensitive data, and hence no data to be leaked.  We stress that this
design also requires at least two physical machines in order to provide these
security guarantees.

@@

As the name suggests, maintenance mode is intended to allow a user to maintain
their VM's software environment, performing actions such as updating software
and uploading any files needed to conduct their research when in secure mode.
In accordance with these needs, maintenance mode allows the user full Internet
access and persistent storage (i.e.  changes to the virtual hard disk will
persist on reboot) in exchange for a lack of access to the sensitive data and
the user's persistent secure storage volume.

When a user initiates the switch to secure mode, their system state is migrated
to a secure machine.  Once this migration has completed, the state of the
maintenance VM is frozen, and remains in this frozen state until the user exits
secure mode.  Upon exit from secure mode, the maintenance VM is simply unfrozen,
and continues processing from the point at which it left off at the time of the
original switch to secure mode.

@@

When the user switches to secure mode, the state of their VM, previously in
maintenance mode, is migrated to a secure machine where it continues execution.
By switching to secure mode, the user sacrifices network access and virtual hard
disk storage persistence in exchange for access to the sensitive data and their
secure storage volume, which is mounted upon entering secure mode.  Any changes
to the normal storage are stored only in the machine's virtual memory, and
discarded when the user exits secure mode.  Only changes to the user's secure
storage volume will persist to the next time the user enters secure mode.
Indeed, when the user switches back to maintenance mode from secure mode, all
accumulated state change after the original switch into secure mode is discarded
from memory, and the maintenance VM simply resumes execution from where it left
off prior to the switch.

@@

A prototype of the Data Capsules system has been developed using the KVM virtual
machine manager, which is built in to the Linux operating system, along with the
user-level counterpart application qemu to achieve virtualization.  A set of
Bash shell scripts were written to implement the basic operations for managing
Data Capsules VMs.  Although the current prototype does not fully support the
migration-based mode change described in this paper, migration-based switching
has been implemented and demonstrated, and the only remaining work is to fully
integrate it into the existing code.

@@

The Data Capsules prototype uses iptables rules to implement a simple set of
firewall rules for VMs.  All VMs on a machine reside on a bridged, private
network.  In secure mode, additional rules are imposed on the private IP address
of the VM, restricting all network access except that between the VM and the IP
address of the one or more servers containing the sensitive data, as written in
a simple configuration file.

@@

One currently unresolved issue with this approach, and any using a mounted disk
volume, is the issue of caching by the operating system of any changes to the
secure volume.  Since changes typically are not immediately written back to a
disk volume, this raises the possibility that a user could switch out of secure
mode before all data is written back, resulting in data loss.  Currently, we
warn users of the prototype before they switch out of secure mode that they must
detatch the secure volume or risk data loss.  In the future, we might include
software on preconfigured VMs that can help users know when it is safe to switch
out of secure mode.

@@

Since users ultimately require some method for releasing the results of their
research, qemu was configured to create a channel through the serial port on the
VM to a UNIX domain socket on the host machine.  A simple daemon was written
which runs on the secure machine to listen on this socket for any results to be
released.  It follows an extremely simple protocol that takes the file name,
file size, and file data and exports them to a database where they can then
undergo a review process.  We discuss a potential approach for automating this
review process for candidate release data in Section \ref{sec:Results Audit}

@@

Although a basic prototype has been implemented and is fully functional, there
are still many potential routes for improvement.  We discuss many of these,
including some we are currently investigating and some that we intend to look
into in the future.

@@

For instance, if an end user's device is compromised, the malicious party could
then install malicious software on the VM while in maintenance mode that is
designed to enter secure mode and subsequently encode all of the private data in
the format of the channel medium (e.g. as an image in the case of remode desktop
protocols like VNC).  The data could then be decoded by additional software
running on the compromised client, or even forwarded for decoding on the
attacker's personal infrastructure.  In the case of a 1024x768 screen resolution
refreshing at 15 times per second, an attacker could leak over 2 GB of data from
a protected repository within one minute, or at the very least saturate the
channel connection with the leaked data.

This example clearly illustrates the imperative need to secure the remote access
channel from these types of attack.  There is also, however, a limit to what can
be done, since the user ultimately must have some way of seeing into the VM that
they are remotely administering.  The first obvious step is to secure the
client-server connection using SSL with certificate-based authentication.  While
this would seem trivial to accomplish, there is an unfortunate lack of support
for such a secure connection in existing widely used VNC clients.  We are
currently in the process of modifying an existing VNC client to support SSL
connections with certificates.  Beyond this initial setback, we expect
implementation to be relatively trivial.

Beyond securing the connection, we are currently investigating methods for
securing the end client device using the Trusted Platform Module (TPM)
\cite{tpm} on an end user's device to remotely attest that the end client is
secure.  We envision the final implementation will consist of a USB thumb drive
containing a barebones Linux operating system with an installed VNC client, and
will use the TPM's features to perform a measured boot of Linux and the VNC
client.  Users will plug in the USB drive to their device and boot into the
secure environment in order to access their Data Capsules VMs.  In this way, we
hope to greatly raise the bar for a malicious attacker who wishes to gain
access to sensitive data, while preserving a reasonably uncumbersome method for
access to the end user.

@@

Although Data Capsules aims to ease secure access to sensitive data by removing
the need to audit data and software being imported into a secure environment, it
does not address how to ease the auditing of results data being released from
that same environment.  This is a challenging issue, since a careful attacker
could use any number of methods to disguise the leaking of large amounts of data
as something that appears entirely innocent.

@@

The second issue with designing a mature migration-based mode switch is that of
speed and performance.  Transferring the state of many machines all running in
maintenance mode could easily strain even a very robust network.  Therefore it
is likely that some form of rate limiting will need to be imposed in order to
relieve network pressure.  The consequence of this is of course that the time it
takes to switch modes will increase.  We therefore plan to investigate to find
the best compromise between the two needs of limiting use of network bandwidth
and maximizing switch performance.

@@

Since the ultimate goal of the Data Capsules system is to improve usability of
remote access to sensitive data, it is natural that performance is an important
measure of its success.  Virtual machines must not suffer any slowdown, and the
new mode switching interface that it introduces must not be inconveniently slow.
Additionally, direct feedback from actual users of our system can serve as a
useful measure of its merits in practice.

@@

Fortunately, virtual machine performance is not significantly affected by Data
Capsules, since it requires no significant changes to qemu or KVM.  Basic VM management functions also run quickly, with starting and stopping and deleting VMs taking only fractions of a second.  VM creation has also been made almost instantatneous to the user through clever use of copy-on-write references to base disk images.  This leaves only the performance of switches between modes.

@@

@@

A user study is another important option for evaluating the undeniably
unfamiliar paradigm of having two machine modes for a single conceptual machine.
Although we have no plans to conduct a formal study in the near future, we are
in the process of implementing a prototype of the system that will be used by
humanities researchers for secure access to copyrighted materials.  After
initial alpha testing by a small number of subjects unfamiliar with the system,
feedback was positive, with no apparent complaints or confusion about how it was
used.  We leave more formal and in depth user studies for future work.

@@

In this paper, we introduced Data Capsules, a system that enables more usable
access to sensitive data repositories by eliminating the need to audit data and
software that are brought in to the secure environment.  Data Capsules presents
users with a maintenance mode for fully customizing their software environment,
and allows users to then import this environment into a secured environment
where they can then perform analysis on sensitive data with limited risk of data
leak.  Data Capsules achieves this mode paradigm by migrating underlying VMs
between machines segregated by their access to the network and to the sensitive
data.  A prototype of the system has been implemented, and initial performance
data and end user feedback are both promising.

>>>>

\title{Data Capsules: A Cloud Service for Secure Remote Access to Sensitive Data}

\author{
%Submission to CCSW 2014
Draft Copy: \today\ \currenttime
}

%\numberofauthors{2} 
%
%\author{
%\alignauthor Alexander Crowell \\
%       \affaddr{Computer Science and Engineering Division}\\
%       \affaddr{University of Michigan}\\
%       \affaddr{Ann Arbor, MI, USA}\\
%       \email{crowella@umich.edu}
%\alignauthor Atul Prakash \\
%       \affaddr{Computer Science and Engineering Division}\\
%       \affaddr{University of Michigan}\\
%       \affaddr{Ann Arbor, MI, USA}\\
%       \email{aprakash@umich.edu}
%}

@@

In this paper, we propose Data Capsules, a system that is designed to simplify
access to sensitive data for analysis by trusted remote users, while
maintaining reasonable guarantees of data security.  Data Capsules uses
virtualization to provide remote users with a fully privileged, but secure
environment into which they can bring arbitrary, and even potentially
malicious, software or data in order to analyze sensitive data, while
minimizing the available channels for data leaks.  Our early implementation
realizes much of this protection, providing a basic framework for secure
analysis of data that addresses many aspects of network, storage, and covert
channel security, and has been adopted for use in the HathiTrust Research
Center.

@@

\keywords{data capsules, data privacy, virtualization, cloud computing}

@@

In the age of the Internet, new data is being created at incredible rates
\cite{digital-universe}.  Along with this wave of new data, new tools have in
turn been created to help analyze and gain insight from it. Unfortunately,
access to data is not always a simple matter.  Some types of data may be
sensitive, including data protected by copyright, data representing the
personal information of individuals, as well as any data that may contain
secrets that its owner does not wish to have revealed, for example corporate or
state secrets.

The current state-of-the-art \cite{vrdc} in systems deployed for the purpose
of enabling access to these types of data use remotely accessed machines, which
can be used by researchers without the need for travel to conduct their
analysis. But beyond this apparent convenience, rules for using such systems
are often highly restrictive, with researchers limited to using predefined
software environments, and being subject to strict auditing requirements for
all software and data that is uploaded.

To solve this substantial inconvenience, we propose a new system which we call
Data Capsules.  While, to an end user, a Data Capsules environment will
minimally differ from the type of environment provided by the aforementioned
state-of-the-art system, its design grants end users privileged access to
install software of their choice --- even down to the virtualization-compatible
operating system of their choice --- without requiring a review process and
without any added theoretical risks to the security of the data being provided.

Indeed, any system handling sensitive data must be carefully designed to close
potential channels for leaking that data as thoroughly as possible.  Our plans
for Data Capsules intend for the system to eventually address threats ranging
from more aggressive attacks, such as attempts to leak data over the network or
other conventional channels, to more subtle attacks, such as an attempt by an
attacker to leak large volumes of sensitive data by surreptitiously exploiting
the trusted user's own remote connection to the machine.  And our current
prototype is already capable of defending against many types of attacks, both
large and small, which we discuss in our description of the system's design.

Data Capsules makes use of virtualization as a key element in achieving this
security and usability goal.  Under the Data Capsules use model, a researcher
would be allocated a virtual machine (henceforth abbreviated as VM, or guest
VM), which they could then prepare and maintain with whatever software and
underlying operating system is best suited to the intended purpose.  Each Data
Capsules VM is endowed with two modes of operation.  In \emph{maintenance}
mode, the researcher is granted full Internet access, which enables them to
easily configure and maintain the software environment almost as if they were
using a their own personal machine; however, in maintenance mode, the
researcher does not have access to the sensitive data.  In order to gain access
to this data, the user must switch their VM to the second mode, which we call
\emph{secure} mode.  In secure mode, the user's VM state is saved, and
execution continues on an alternate machine that is firewalled to prevent
Internet access, but which does have access to the sensitive data.

@@

When a user wishes to further change their software environment in order to
continue data analysis, they may then switch back to maintenance mode, after
which the secure storage volume is detached and all VM state since the
previous switch to secure mode is discarded, with the VM continuing to run from
the point it left off at when last in maintenance mode.  In this way, every
trace of the operations performed on the sensitive data is erased and forgotten.

Furthermore, assuming all sensitive data is read-only, and is properly backed
up, Data Capsules functionality means that there is no need to audit installed
software for malware, as even infected software would be unable to exfiltrate
the data from the secure mode environment.

@@

While this system provides access to Census data, and seems robust to data
leaks, it is also highly restrictive, requiring researchers to relocate and use
foreign hardware, and potentially software as well, to conduct their research.

@@

However, this state-of-the-art solution still suffers from significant
drawbacks, in that researchers are largely limited to using the preallocated
environments supplied by the VRDCs.  In order to upload software or data not
already available in a virtual desktop, it must first undergo a manual review
process.  For large software suites, this can at best add significant
inconvenience, and at worst expose the system to a malicious attack, as there
is no known method for reliably detecting malicious software.  Data Capsules is
designed so that no review process is necessary to upload software or data to
an environment, while providing strong guarantees of data privacy by design.

@@

QubesOS \cite{qubes} is a system for isolating a user's desktop environment
into secure compartments, called ``Qubes," which are intended to prevent
malicious code from compromising the entire system by restricting its effect to
the compartment in which it resides.  Although Qubes is very effective for this
purpose, it would not serve as a particularly effective tool for data access,
as it suffers from the same challenge of how to move data from one compartment,
which is presumably unrestricted, into another compartment containing the
sensitive data to be operated on.  Data Capsules offers a clear design for
allowing the import of data and code into an environment that can still offer
secure access to sensitive data.

Storage Capsules \cite{capsules-borders} is a virtualization-based system for
protecting data privacy on individual machines.  Under Storage Capsules, a
machine runs a virtual machine manager (VMM) system, which in turn runs the
user's operating system environment within a VM.  On the same machine, there is
a second VM running an interface that serves as a ``storage capsule" for
sensitive data.  The user's main VM is endowed with two modes of operation,
\emph{normal} and \emph{secure} mode, and similar to Data Capsules, normal mode
grants full network privileges while secure mode grants access to the sensitive
data, via communication with the storage capsule VM.

@@

Specialized operating systems that enforce information flow control, such as
HiStar \cite{histar}, offer low-level operating system primitives that allow
developers to design programs in such a way that data leaks are not possible.
As with SELinux, there is still a substantial burden of having to craft policies
that prevent these malicious data flows.  Additionally, programs must be
specifically developed for compatibility with HiStar, creating an even greater
burden on users so long as that system is not widely supported by the developer
community.

Another potential approach to controlling information flow in databases is that
of differential privacy \cite{differential, differential-survey}. This uses
statistical methods to ensure that information about individual entries in a
database is not revealed, while still allowing statistical analysis of the
underlying data.  While differential privacy is indeed promising, this approach
clearly has very limited applications, as it only applies to highly structured
data with obvious approaches for statistical analysis.  Data Capsules allows
analysis of all types of data, structured or unstructured, with all types of
software, while maintaining reasonable guarantees of security.

@@

The choice of virtualization versus more lightweight containment approaches,
such as Linux containers \cite{lxc}, is based both on this greater
trustworthiness of a smaller hypervisor versus having to trust an entire
operating system kernel, as well as the benefit of users having greater freedom
to choose their own virtualization-compatible operating system, rather than
having a single software environment chosen for them.

@@

The major element that is used to isolate VMs that are in the middle of
accessing sensitive data is live migration \cite{migration-clark}.  When a VM
switches from maintenance mode to secure mode, the underlying virtual machine
manager migrates that VM to a separate physical machine that has access to the
sensitive data, but which is configured to lack network access for any purpose
other than to receive other incoming VM migrations and access read-only VM disk
images that are served over a network-based file system.  In other words, VMs
in maintenance mode are strictly segregated within separate physical machines
from VMs running in secure mode.  We can therefore refer to a physical machine
running only maintenance VMs as a \emph{maintenance machine}, and likewise a
\emph{secure machine} runs only secure mode instantiations of VMs.  We
illustrate an example deployment in Figure \ref{fig:configuration}.

By adopting this segregated design, the threat of data leak via covert channels
within a single physical machine is eliminated, so long as the trusted
computing base remains uncompromised.  Intermachine covert channels along the
connection between maintenance machines and secure machines are the greatest
threat.  However, this threat is relatively low, since data largely travels one
way from maintenance machines to secure machines.  We plan to consider more
robust, active protections to this channel as future work.  The remaining risk
is that of one VM in secure mode leaking data to another VM that is also in
secure mode, but there is no reason to want to do so, as both VMs would already
have access to the entire secure repository.  On a machine running VMs in
maintenance mode, there is no way to access the sensitive data, and hence no
data to be leaked.  We stress that this design also requires at least two
physical machines in order to provide these security guarantees.

@@

\caption{An example deployment configuration for a Data Capsules system.
Dotted boxes denote physical separation, which is necessary between maintenance
machines, secure machines, and the remote access proxy.  The sensitive data
server may exist separately, or it may reside on one of the secure machines.
The only connections between maintenance and secure machines will be for the
purpose of transferring virtual machine state to secure machines, and for
read-only accesses to the VM disk images.  This data largely travels in one
direction, from maintenance machines to secure machines.}

@@

@@

As the name suggests, maintenance mode is intended to allow a user to maintain
their VM's software environment, performing actions such as updating software
and uploading any files needed to conduct their research when in secure mode.
In accordance with these needs, maintenance mode allows the user full Internet
access and persistent storage (i.e.  changes to the virtual hard disk will
persist across reboots and mode changes) in exchange for a lack of access to
the sensitive data and the user's persistent secure storage volume.

As a user works in maintenance mode, the system state of their VM is
synchronized to a secure machine in anticipation of an eventual switch to
secure mode.  When a user initiates the switch to secure mode, the migration of
system state to the secure machine is completed, the state of the maintenance
VM is frozen, and it remains in this frozen state until the user exits secure
mode.  Upon exit from secure mode, the maintenance VM is simply unfrozen, and
continues processing from the point at which it left off at the time of the
original switch to secure mode.

@@

When the user switches to secure mode, the state of their VM, previously in
maintenance mode, is migrated to a secure machine where it continues execution.
By switching to secure mode, the user sacrifices network access and virtual
hard disk storage persistence in exchange for access to the sensitive data and
their secure storage volume, which is mounted upon entering secure mode.  Any
changes to the normal storage are stored only in the secure machine's virtual
memory, and discarded when the user exits secure mode.  Only changes to the
user's secure storage volume will persist to the next time the user enters
secure mode.  Indeed, when the user switches back to maintenance mode from
secure mode, all accumulated state change after the original switch into secure
mode is discarded from memory, and the maintenance VM simply resumes execution
from where it left off prior to the switch.

@@

A prototype of the Data Capsules system has been developed using the KVM
virtual machine hypervisor, which is built in to the Linux operating system,
along with the user-level counterpart application qemu to achieve
virtualization.  A set of Bash shell scripts were written to implement the
basic operations for managing Data Capsules VMs.  Although the current
prototype does not fully support the migration-based mode change described in
this paper, migration-based switching has been implemented and demonstrated,
and the only remaining work is to fully integrate it into the existing code.

@@

The Data Capsules prototype uses iptables rules to implement a simple set of
firewall rules for VMs.  All VMs on a machine reside on a bridged, private
network.  In secure mode, additional rules are imposed on the private IP
address(es) of the VM, restricting all network access except that between the
VM and the IP address of the one or more servers containing the sensitive data,
as written in a simple configuration file.

@@

One currently unresolved issue with this approach, and any using a mounted disk
volume, is the issue of caching by the operating system of any changes to the
secure volume.  Since changes typically are not immediately written back to a
disk volume, this raises the possibility that a user could switch out of secure
mode before all data is written back, resulting in data loss.  Currently, we
warn users of the prototype before they switch out of secure mode that they
must first detach the secure volume or risk data loss.  In the future, we might
include software on popular preconfigured VMs that can help users know when it
is safe to switch out of secure mode.

@@

Since users ultimately require some method for releasing the results of their
research, qemu was configured to create a channel through the serial port on
the VM to a UNIX domain socket on the host machine.  A simple daemon was
written which runs on the secure machine to listen on this socket for any
results to be released.  It follows an extremely simple protocol that takes the
file name, file size, and file data and exports them to a database where they
can then undergo a review process.  We discuss a potential approach for
partially automating this review process for candidate release data in Section
\ref{sec:Results Audit}.

@@

Although a basic prototype has been implemented and is fully functional, there
are still many potential routes for improvement.  We discuss several of these,
including some that we are currently investigating and some that we intend to
look into in the future.

@@

For instance, if an end user's device is compromised, the malicious party could
then install malicious software on the VM while in maintenance mode that is
designed to enter secure mode and subsequently encode all of the private data in
the format of the channel medium (e.g. as an image in the case of remote desktop
protocols like VNC).  The data could then be decoded by additional software
running on the compromised client, or even forwarded for decoding on the
attacker's personal infrastructure.  In the case of a 1024x768 screen resolution
refreshing at 15 times per second, an attacker could leak over 2 GB of data from
a protected repository within one minute, or at the very least saturate the
channel connection with the leaked data.

This example clearly illustrates the imperative need to secure the remote
access channel from these types of attacks.  There is also, however, a limit to
what can be done, since the user ultimately must have some way of seeing into
the VM that they are remotely administering.  The first obvious step is to
secure the client-server connection using SSL with certificate-based
authentication.  While this would seem trivial to accomplish, there is an
unfortunate lack of support for such a secure connection in existing widely
used VNC clients.  We are currently in the process of modifying an existing VNC
client to support SSL connections with certificates.  Beyond this initial
setback, we expect implementation to be relatively trivial.

Beyond securing the connection, we are currently investigating methods for
securing the end client device using the Trusted Platform Module (TPM)
\cite{tpm} on an end user's device to remotely attest that the end client is
secure.  We envision the final implementation consisting of a USB thumb drive
containing a barebones Linux operating system with an installed VNC client,
which will use the TPM's features to perform a measured boot of Linux and the
VNC client.  Users will plug in the USB drive to their device and boot into the
secure environment in order to access their Data Capsules VMs.  In this way, we
hope to greatly raise the bar for a malicious attacker who wishes to gain
access to sensitive data by attacking the remote access connection, while
preserving a reasonably uncumbersome method for access to the end user.

@@

Although Data Capsules aims to ease secure access to sensitive data by removing
the need to audit data and software being imported into a secure environment,
it does not address how to ease the auditing of results data being released
from that same environment.  This is a challenging issue, since a careful
attacker could potentially use any number of methods to disguise the leaking of
large amounts of data as something that appears entirely innocent.

@@

The second issue with designing a mature migration-based mode switch is that of
speed and performance.  Transferring the state of many machines all
simultaneously running in maintenance mode could easily strain even a very
robust network.  Therefore, it is likely that some form of rate limiting will
need to be imposed in order to relieve network pressure.  The consequence of
this is of course that the time it takes to switch modes will increase.  We
therefore plan to investigate to find the best compromise between the two needs
of limiting use of network bandwidth and maximizing switch performance.

\subsection{Utilizing Excess Computational Power}

One important design choice made in the Data Capsules system is the segregation
of VMs running in maintenance mode from those running in secure mode.  This
choice adds robustness to some types of threats, but in exchange requires
double the computational power to realize an actual deployment.  Since VMs can
only exist in at most one mode at a time, there is a high potential for wasted
computational power in practice.  While this is not as much of an issue for
maintenance machines, which face no significant restrictions, besides not having
access to the sensitive data, and hence could put their resources to any
desired use, secure machines are much more complicated.

While we leave the investigation into how secured computational resources could
be used most efficiently as future work, we do stress that the cost of
deploying a Data Capsules-enabled system is not strictly double the cost of an
equivalent non-segregated system, since computational resources on maintenance
machines can be used to offset these additional costs without increased
security risk.

@@

Since the ultimate goal of the Data Capsules system is to simplify usability of
remote access to sensitive data, it is natural that performance is an important
measure of its success.  Virtual machines must not suffer any slowdown, and the
mode switching interface that it introduces must not be inconveniently slow.
Additionally, direct feedback from actual users of our system can serve as a
useful measure of its merits in practice.

@@

Fortunately, virtual machine performance is not significantly affected by Data
Capsules, since it requires no significant changes to qemu or KVM.  Basic VM
management functions also run quickly, with starting and stopping and deleting
VMs taking only fractions of a second.  VM creation has also been made almost
instantaneous to the user through clever use of copy-on-write references to base
disk images.  This leaves the functional performance of switches between modes.

@@

One final issue that requires investigation is memory use in secure mode.  Since
system state in this mode is not persistent, changes to it are stored in virtual
memory.  For a more comprehensive evaluation, we intend to investigate this
memory usage, and if necessary find ways to reduce it or any effects it may have
on long term performance.

@@

A user study is another important option for evaluating the undeniably
unfamiliar paradigm of having two machine modes for a single conceptual
machine.  Although we have no plans to conduct a formal study in the near
future, we are in the process of implementing a prototype of the system that
will be used by humanities researchers for secure access to copyrighted
material in the HathiTrust Research Center.  After initial alpha testing by a
small number of subjects unfamiliar with the system, feedback was positive,
with no apparent complaints or confusion about how it was used.  We leave more
formal and in depth user studies for future work.

@@

In this paper, we introduced Data Capsules, a system that enables simplified
end-user access to sensitive data repositories by eliminating the need to audit
data and software that are brought in to the secure environment.  Data Capsules
presents users with a maintenance mode for fully customizing their software
environment, and allows users to then import this environment into a secured
domain where they can then perform analysis on sensitive data with limited risk
of data leak.  Data Capsules achieves this mode paradigm by migrating
underlying VMs between machines segregated by their access to the network and
to the sensitive data.  A prototype of the system has been implemented with
limited deployment to test users, and initial performance data and end user
feedback are both promising.
